---
title: "ST-661 - R for data Analytics Mini Project"
author: "Atikant Negi"
output: 
  html_document :
  code_folding : hide
---

```{r setup, include=FALSE}
library(pxR)
knitr::opts_chunk$set(echo = TRUE)
```


## Road accident casualties in Ireland{.tabset}
### Synopsis
__Objective__ : The objective of the project is to go through the road accident casualty data in Ireland from year 2005 - 2018 and answer some questions around it like what gender, age groups and road type users are more likely to encounter a casualty. What are the trends over the years around the injuries and people killed in road accidents. Predicting the outcomes of the near future events based on the previous data.

---

__Solution__ : Going through and analyzing the data using various pairs of explanatory variables we were able to draw conclusion. We used a variety of packages in R to draw the results.

---

__Insights__ : Various insights were drawn though the analysis such as:\

* Most no. of cases(9442) came in year 2009.
* Most no. of deaths (388) were reported in year 2005. 
* The no. of deaths have been gradually decreasing although the average number of cases (8017) reported remains approximately constant.

---


### Packages Requirement

The following packages were used in the project:

* pxR
* ggplot2
* dplyr
* relaimpo
* Knitr
* Fitdistrplus
* car

```{r ,eval=FALSE,}
library(pxR)
library(ggplot2)
library(relaimpo)
library(car)
```


### Data Preparation {.tabset}

This sections consist of all the steps taken in getting the data analysis ready. The steps are explained in detail below.

#### Data Import

__Source of Data__ : https://statbank.cso.ie/px/pxeirestat/Database/eirestat/Road%20Safety%20Statistics/Road%20Safety%20Statistics_statbank.asp?SP=Road%20Safety%20Statistics&Planguage=0&ProductID=DB_RSA
\

__File Name__ : ROA16
\

__File Type__ : .Px
\

__Import command__ : 
\

```{r}
options(width = 120)
data<-as.data.frame(read.px("ROA16.px"))
head(data)
colnames(data)
dim(data)
```
__Data Cleaning__ : 
The data is flawless and needs no data cleaning. The values seems in line and there are no missing or invalid values in the data-set.


#### Data Preview

#### Data Description

There are 5 categorical data in the data-set i.e. 
\

* Statistics has 3 levels
```{r}
levels(data$Statistic)
```


* Years has 14 levels as:
```{r}
levels(data$Year)
```

* Road user type has 8 levels
```{r}
levels(data$Road.User.Type)
```

* Sex has 2 levels
```{r}
levels(data$Sex)
```

* age group has 13 levels
```{r}
levels(data$Age.Group)
```




### Data Anlaysis
### Summary