---
title: "ST-661 - R for data Analytics Mini Project"
author: "Atikant Negi"
output: 
  html_document :
    code_folding: "hide"
---

```{r setup, include=FALSE}
library(pxR)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE,fig.width = 11,fig.height = 8)
```


## Road accident casualties in Ireland{.tabset}
### Synopsis
__Objective__ : The objective of the project is to go through the road accident casualty data in Ireland from year 2005 - 2018 and answer some questions around it like what gender, age groups and road type users are more likely to encounter a casualty. What are the trends over the years around the injuries and people killed in road accidents. Predicting the outcomes of the near future events based on the previous data.

---

__Solution__ : Going through and analyzing the data using various pairs of explanatory variables we were able to draw conclusion. We used a variety of packages in R to draw the results.

---

__Insights__ : Various insights were drawn though the analysis such as:\

* Most no. of cases(9442) came in year 2009.
* Most no. of deaths (388) were reported in year 2005. 
* The no. of deaths have been gradually decreasing although the average number of cases (8017) reported remains approximately constant.
* The most number of casualties are seen in car drivers with its peak in 2009 and the lowest in 2013.
* 

---


### Packages Requirement

The following packages were used in the project:

* pxR
* ggplot2
* dplyr
* relaimpo
* Knitr
* Fitdistrplus
* car

```{r ,eval=FALSE,}
install.packages("ggplot2")
library(pxR)

library(ggplot2)
library(relaimpo)
library(car)
```


### Data Preparation {.tabset}

This sections consist of all the steps taken in getting the data analysis ready. The steps are explained in detail below.

#### Data Import

__Source of Data__ : https://statbank.cso.ie/px/pxeirestat/Database/eirestat/Road%20Safety%20Statistics/Road%20Safety%20Statistics_statbank.asp?SP=Road%20Safety%20Statistics&Planguage=0&ProductID=DB_RSA
\

__File Name__ : ROA16
\

__File Type__ : .Px
\

__Import command__ : 
\

```{r}
options(width = 120)
data<-as.data.frame(read.px("ROA16.px"))
head(data)
colnames(data)
dim(data)
```
__Data Cleaning__ : 
The data is flawless and needs no data cleaning. The values seems in line and there are no missing or invalid values in the data-set.


#### Data Preview

```{r}
options(width = 120)
head(data)
```

#### Data Description

There are 5 categorical data in the data-set i.e. 
\

* Statistics has 3 levels
```{r}
levels(data$Statistic)
```


* Years has 14 levels as:
```{r}
levels(data$Year)
```

* Road user type has 8 levels
```{r}
levels(data$Road.User.Type)
```

* Sex has 2 levels
```{r}
levels(data$Sex)
```

* age group has 13 levels
```{r}
levels(data$Age.Group)
```




### Data Anlaysis {.tabset}

#### General analysis

```{r}
# if we see the total no. of casualties over the years we can sum it as the below data set
d1<-subset(data,Statistic=="All Killed and Injured Casualties (Number)"
           &
             Road.User.Type=="All road users" 
           & 
             Age.Group=="All ages"
           & 
             Sex=="Male")%>%
  group_by(Year)

d2<-subset(data,Statistic=="All Killed and Injured Casualties (Number)"
           &
             Road.User.Type=="All road users" 
           & 
             Age.Group=="All ages"
           & 
             Sex=="Female")%>%
  group_by(Year)
total_casuality<-as_tibble(full_join(d1,d2,by=c("Year","Road.User.Type","Age.Group")))
total_casuality<-mutate(total_casuality,Total_vaue=total_casuality$value.x+total_casuality$value.y)


# if we see the total no. of injuries over the years we can sum it as the below data set
d_only_injured_male<-subset(data,Statistic=="Injured Cassualties (Number)"
           &
             Road.User.Type=="All road users" 
           & 
             Age.Group=="All ages"
           & 
             Sex=="Male")%>%
  group_by(Year)

d_only_injured_female<-subset(data,Statistic=="Injured Cassualties (Number)"
           &
             Road.User.Type=="All road users" 
           & 
             Age.Group=="All ages"
           & 
             Sex=="Female")%>%
  group_by(Year)
total_casuality_injured<-as_tibble(full_join(d_only_injured_male,d_only_injured_female,by=c("Year","Road.User.Type","Age.Group")))
total_casuality_injured<-mutate(total_casuality_injured,Total_vaue_injured=total_casuality_injured$value.x+total_casuality_injured$value.y)

  
normalized_data <-data.frame("year"=c(2005:2018),
                             "Road.User.Type"=rep("All road users",14),
                            "Injured"=total_casuality_injured$Total_vaue_injured,
                             "Killed_and_Injured"=total_casuality$Total_vaue,
                            "killed"=total_casuality$Total_vaue-total_casuality_injured$Total_vaue_injured)

normalized_data
# summary for casualties
summary(normalized_data)

# plot for the total casualties
plot(normalized_data$year,normalized_data$Killed_and_Injured,col="green",type="l",xlab="Year",ylab = "Numbers",ylim=c(0,10000))
lines(normalized_data$year,normalized_data$Injured,col="blue",type="l" )
lines(normalized_data$year,normalized_data$killed,col="red",type="l")

legend(2015, 10500, legend=c("Killed", "Injured","Total casualties"),
       col=c("red", "blue" ,"green"), lty=1, cex=0.75)


Male_female<-data.frame("year"=c(2005:2018),
                        "male_injured"=d_only_injured_male$value,
                        "female_injured"=d_only_injured_female$value,
                        "male_killed"=d1$value-d_only_injured_male$value,
                        "female_killed"=d2$value-d_only_injured_female$value
                          )
plot(Male_female$year,Male_female$female_injured,col="magenta",type="l",xlab="Year",ylab = "Numbers",ylim=c(0,6000))
lines(Male_female$year,Male_female$male_injured ,col="blue",type="l" )
lines(Male_female$year,Male_female$male_killed,col="cyan",type="l")
lines(Male_female$year,Male_female$female_killed,col="pink",type="l")

legend(2015, 6250, legend=c("Deceased Male", "Injured Male","Deceased Female", "Injured female"),
       col=c("cyan", "blue" ,"pink","magenta"), lty=1, cex=0.75)

```
\
\

#### Road user analysis

```{r}
options(width = 140)
library(ggplot2)
roadUser<-subset(data,Statistic=='All Killed and Injured Casualties (Number)' &
                (
                  Road.User.Type=='Pedestrians' |
                  Road.User.Type=='Pedal cyclists' |
                  Road.User.Type=='Motor cyclists' |
                  Road.User.Type=='Car drivers' |
                  Road.User.Type=='Car passengers' |
                  Road.User.Type=='Other road users'
                    )
              & Age.Group=="All ages")

roadUser_normalized_data<-data.frame("year"=roadUser$Year,"UserType"=roadUser$Road.User.Type,"Sex"=roadUser$Sex,"value"=roadUser$value)



ggplot(roadUser_normalized_data,aes(x=year,y=value,col=UserType,group=UserType))+
  geom_point()+
  geom_line()+
  facet_wrap(.~Sex)

```
\
\

#### Age group analysis

```{r}
AgeGroup<-subset(data,Statistic=='All Killed and Injured Casualties (Number)' &
                      Age.Group !="All ages" 
             & Road.User.Type=="All road users")



ggplot(AgeGroup,aes(x=Year,y=value,col=Age.Group,group=Age.Group))+
  geom_point()+
  geom_line()+
  facet_wrap(.~Sex)

```




### Summary