---
title: "R_PROJECT"
author: "AVANTIKA SINGH (20250339)"
date: "12/6/2020"
output: html_document
---


# Determining the SAFESR ROAD USER TYPE based on AGE GROUP AND GENDER BASED CASUALTIES ANALYSIS OVER TIME PERIOD OF 14 YEARS.

# Categorical Aggregate of Total Casualties For Each ROad User Type for 14 years.

```{r}
library(pxR)
df = as.data.frame(read.px("/Users/macbp/Downloads/ROA16.px"))
names(df)
dim(df)
head(df)
```

```{r}
p<-ggplot(data=df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type != "All road users" & df$Road.User.Type != "All Car users", ], aes(x=Road.User.Type, y=value)) +
  geom_bar(stat="identity")
p
```


```{r}
sub1 = df[df$Statistic !='All Killed and Injured Casualties (Number)' & df$Road.User.Type != "All road users" & df$Road.User.Type != "All Car users", ]
ggplot(data=sub1, aes(x=Road.User.Type, y=value, fill=Statistic), position = "fill") +
  geom_bar(stat="identity")
```

# Car Drivers having the highest Casualties Count.
# Deep Dicing in to Car Driver Casualties for Gender and AGe Group.

# Proportion of Male and Female Car Driving Casualties Trend over 14 years.

```{r}
sub1 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Sex == 'Female', ]
temp1 <- aggregate(value~Year, sub1, sum)
sub2 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "All road users" & df$Sex == 'Female', ]
temp2 <- aggregate(value~Year, sub2, sum)
Female_prop = round(temp1$value/temp2$value, 2)
Female_prop

sub3 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Sex == 'Male', ]
temp3 <- aggregate(value~Year, sub3, sum)
sub4 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "All road users" & df$Sex == 'Male', ]
temp4 <- aggregate(value~Year, sub4, sum)
Male_prop = round(temp3$value/temp4$value, 2)
Male_prop
```


# Proportion of Male and Female Car Drivind Death Casualties Trend over 14 years.

```{r}
sub1 = df[df$Statistic =='Killed Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Sex == 'Female', ]
temp1 <- aggregate(value~Year, sub1, sum)
sub2 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Sex == 'Female', ]
temp2 <- aggregate(value~Year, sub2, sum)
Female_kld_prop = round(temp1$value/temp2$value, 2)
Female_kld_prop

sub3 = df[df$Statistic =='Killed Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Sex == 'Male', ]
temp3 <- aggregate(value~Year, sub3, sum)
sub4 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Sex == 'Male', ]
temp4 <- aggregate(value~Year, sub4, sum)
Male_kld_prop = round(temp3$value/temp4$value, 2)
Male_kld_prop
```

# Categorical Count of AGe Group WIse Casualties for 14 years.

```{r}
sub1 = df[df$Statistic !='All Killed and Injured Casualties (Number)' & df$Road.User.Type != "All road users" & df$Road.User.Type != "All Car users" & df$Age.Group != "All ages", ]
ggplot(data=sub1, aes(x=Age.Group, y=value, fill=Statistic), position = "fill") +
  geom_bar(stat="identity")
```


# 25-34 years AGE GROUP Has highest Casualties COunt.

# Analyzing behaviour of Top 4 Casualties Counts AGe Groups Based on Car DRIVING CASUALTY PROPORTION AND KILLED Car DRIVING CASUALTY PROPORTION TRENDS.

```{r}
sub11 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & 
            (df$Age.Group == '15 - 17 years' | df$Age.Group == '18 - 20 years' | df$Age.Group == '21 - 24 years'), ]
temp11 <- aggregate(value~Year, sub11, sum)
sub12 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "All road users" & 
            (df$Age.Group == '15 - 17 years' | df$Age.Group == '18 - 20 years' | df$Age.Group == '21 - 24 years'), ]
temp12 <- aggregate(value~Year, sub12, sum)
age1_prop = round(temp11$value/temp12$value, 2)
age1_prop

sub21 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Age.Group == '25 - 34 years', ]
temp21 <- aggregate(value~Year, sub21, sum)
sub22 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "All road users" & df$Age.Group == '25 - 34 years', ]
temp22 <- aggregate(value~Year, sub22, sum)
age2_prop = round(temp21$value/temp22$value, 2)
age2_prop

sub3 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Age.Group == '35 - 44 years', ]
temp3 <- aggregate(value~Year, sub3, sum)
sub4 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "All road users" & df$Age.Group == '35 - 44 years', ]
temp4 <- aggregate(value~Year, sub4, sum)
age2_prop = round(temp3$value/temp4$value, 2)
age2_prop

sub3 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers" & df$Age.Group == '45 - 54 years', ]
temp3 <- aggregate(value~Year, sub3, sum)
sub4 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "All road users" & df$Age.Group == '45 - 54 years', ]
temp4 <- aggregate(value~Year, sub4, sum)
age3_prop = round(temp3$value/temp4$value, 2)
age3_prop
```

# Overall Trend of Casualties and Killed Casualties Proportions

```{r}
sub1 = df[df$Statistic =='Killed Casualties (Number)', ]
temp1 <- aggregate(value~Year, sub1, sum)
sub2 = df[df$Statistic =='Injured Cassualties (Number)', ]
temp2 <- aggregate(value~Year, sub2, sum)
sub3 = df[df$Statistic =='All Killed and Injured Casualties (Number)', ]
temp3 <- aggregate(value~Year, sub3, sum)
Killed_prop = round(temp1$value/temp3$value, 2)
Injured_prop = round(temp2$value/temp3$value, 2)
Injured_prop
Killed_prop
```

# Road User TYpe Casualties Proprtion Trend

```{r}
sub1 = df[df$Road.User.Type =='Pedestrians' & df$Statistic == 'All Killed and Injured Casualties (Number)', ]
temp1 <- aggregate(value~Year, sub1, sum)
sub2 = df[df$Road.User.Type =='Pedal cyclists' & df$Statistic == 'All Killed and Injured Casualties (Number)', ]
temp2 <- aggregate(value~Year, sub2, sum)
sub3 = df[df$Road.User.Type =='Motor cyclists' & df$Statistic == 'All Killed and Injured Casualties (Number)', ]
temp3 <- aggregate(value~Year, sub3, sum)
sub4 = df[df$Road.User.Type =='Car drivers' & df$Statistic == 'All Killed and Injured Casualties (Number)', ]
temp4 <- aggregate(value~Year, sub4, sum)
sub5 = df[df$Road.User.Type =='Car passengers' & df$Statistic == 'All Killed and Injured Casualties (Number)', ]
temp5 <- aggregate(value~Year, sub5, sum)
sub6 = df[df$Road.User.Type =='Other road users' & df$Statistic == 'All Killed and Injured Casualties (Number)', ]
temp6 <- aggregate(value~Year, sub6, sum)
sub7 = df[df$Road.User.Type =='All road users' & df$Statistic == 'All Killed and Injured Casualties (Number)', ]
temp7 <- aggregate(value~Year, sub7, sum)
Pedestrian_prop = round(temp1$value/temp7$value, 2)
Pcyclist_prop = round(temp2$value/temp7$value, 2)
Mcyclist_prop = round(temp3$value/temp7$value, 2)
CarD_prop = round(temp4$value/temp7$value, 2)
CarP_prop = round(temp5$value/temp7$value, 2)
Pedestrian_prop
Pcyclist_prop
Mcyclist_prop
CarD_prop
CarP_prop
```


# Road User TYpe Killed Casualties Proprtion Trend

```{r}
sub11 = df[df$Statistic =='Killed Casualties (Number)' & df$Road.User.Type == "Pedestrians", ]
temp11 = aggregate(value~Year, sub11, sum)
sub12 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Pedestrians", ]
temp12 = aggregate(value~Year, sub12, sum)
sub21 = df[df$Statistic =='Killed Casualties (Number)' & df$Road.User.Type == "Pedal cyclists", ]
temp21 = aggregate(value~Year, sub21, sum)
sub22 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Pedal cyclists", ]
temp22 = aggregate(value~Year, sub2, sum)
sub31 = df[df$Statistic =='Killed Casualties (Number)' & df$Road.User.Type == "Motor cyclists", ]
temp31 = aggregate(value~Year, sub31, sum)
sub32 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Motor cyclists", ]
temp32 = aggregate(value~Year, sub32, sum)
sub41 = df[df$Statistic =='Killed Casualties (Number)' & df$Road.User.Type == "Car drivers", ]
temp41 = aggregate(value~Year, sub41, sum)
sub42 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car drivers", ]
temp42 = aggregate(value~Year, sub42, sum)
sub51 = df[df$Statistic =='Killed Casualties (Number)' & df$Road.User.Type == "Car passengers", ]
temp51 = aggregate(value~Year, sub51, sum)
sub52 = df[df$Statistic =='All Killed and Injured Casualties (Number)' & df$Road.User.Type == "Car passengers", ]
temp52 = aggregate(value~Year, sub52, sum)
Ped_kld_prop = round(temp11$value/temp12$value, 2)
PCyc_kld_prop = round(temp21$value/temp22$value, 2)
MCyc_kld_prop = round(temp31$value/temp32$value, 2)
CarD_kld_prop = round(temp41$value/temp42$value, 2)
CarP_kld_prop  = round(temp51$value/temp52$value, 2)
Ped_kld_prop
PCyc_kld_prop
MCyc_kld_prop
CarD_kld_prop
CarP_kld_prop
```



# Hence we conclude that even though the total number of Casualties for the Road User Type - Pedal Cyclist seems to display an increasing trend over the years; yet it has the least proportion of Killed Casualties amongst all other RUT and it is also displaying a slightly decreasing trend.